---
title: "Lab 04: Multinomial Logistic Regression"
subtitle: "Lab 4"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning=FALSE, 
                      message=FALSE,
                      eval=FALSE)
```

## Instructions

-   If you are fitting a model, display the model output in a neatly formatted table. (The `tidy` and `kable` functions can help!).

-   If you are creating a plot, use clear labels for all axes, titles, etc.

-   Commit and push your work to GitHub regularly, at least after each exercise. Write short and informative commit messages.

-   When you're done, we should be able to knit the final version of the QMD in your GitHub as a HTML.

## Data

The data for this assignment comes from an online Ipsos survey that was conducted for the FiveThirtyEight article ["Why Many Americans Don't Vote"](https://projects.fivethirtyeight.com/non-voters-poll-2020-election/). You can read more about the survey design and respondents in the README of the [GitHub repo](https://github.com/fivethirtyeight/data/tree/master/non-voters) for the data.

Respondents were asked a variety of questions about their political beliefs, thoughts on multiple issues, and voting behavior. We will focus on using the demographic variables and someone's party identification to understand whether a person is a probable voter.

The variables we'll focus on are (definitions from the codebook in data set GitHub repo):

-   `ppage`: Age of respondent

-   `educ`: Highest educational attainment category.

-   `race`: Race of respondent, census categories. Note: all categories except Hispanic are non-Hispanic.

-   `gender`: Gender of respondent

-   `income_cat`: Household income category of respondent

-   `Q30`: Response to the question "Generally speaking, do you think of yourself as a..."

    -   1: Republican
    -   2: Democrat
    -   3: Independent
    -   4: Another party, please specify
    -   5: No preference
    -   -1: No response

-   `voter_category`: past voting behavior:

    -   **always**: respondent voted in all or all-but-one of the elections they were eligible in
    -   **sporadic**: respondent voted in at least two, but fewer than all-but-one of the elections they were eligible in
    -   **rarely/never**: respondent voted in 0 or 1 of the elections they were eligible in

You can read in the data directly from the GitHub repo:

```{r}
#load packages
library(tidyverse) #  ggplot and data wrangling
library(nnet) # multinom
library(emmeans) # contrasts
library(ggeffects) # predicted probs
library(knitr) # nice tables
library(patchwork) # nice figures
```

```{r}
voter_data <- read_csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/non-voters/nonvoters_data.csv")
```

```{}
```

# Lab

-   The variable `Q30` contains the respondent's political party identification. Make a new variable that simplifies `Q30` into four categories: "Democrat", "Republican", "Independent", "Other" ("Other" also includes respondents who did not answer the question).

```{r}



```

-   The variable `voter_category` identifies the respondent's past voter behavior. Relevel the variable to make rarely/never the baseline level, followed by sporadic, then always

    ```{r}


    ```

-   Mean center age

    ```{r}

    ```

-   In the [FiveThirtyEight article](https://projects.fivethirtyeight.com/non-voters-poll-2020-election/), the authors include visualizations of the relationship between the voter category and demographic variables such as race, age, education, etc. Select two demographic variables. For each variable, try to replicate the visualizations and interpret the plot to describe its relationship with voter category. Have fun with it: https://www.mikelee.co/posts/2020-02-08-recreate-fivethirtyeight-chicklet-stacked-bar-chart-in-ggplot2.

```{r}

#fig1


```

```{r}

#fig2
```

```{r}
#Combine these plots with patchwork

```

-   Fit a model using mean-centered age, race, gender, income, and education to predict voter category. Show the code used to fit the model, but do **not** display the model output.

    ```{r}


    ```

-   *Should party identification be added to the model?* Include the output from the test, and the conclusion in the context of the data.

    ```{r}



    ```

-   Run a model with `age` `political orientation` and `education` (no interaction).

    ```{r}



    ```

    -   Follow up each term looking at the ***odds of voting sporadically versus rarely/never***

        -   Interpret the intercept in the context of the data

            ```{r}

            ```

        -   Interpret the effect of age in the context of the data

            ```{r}

            ```

        -   Interpret the effect of party ID in the context of the data. Include discussion about which level(s) differ from the baseline.

            ```{r}

            ```

        -   Interpret the effect of education in the context of the data. Include discussion about which level(s) differ from the baseline.

            ```{r}

            ```

    -   Next, plot the predicted probabilities of voter category for `age` and `Party ID` and `education`

        ```{r}

        ```

        ```{r}

        ```

    ```{r}

    ```

-   Write up the model with `age` `political orientation`M and `education`. Make sure you include the $\chi^2$ for each predictor as well as the model fit. Also make sure your write-up includes a discussion of which level(s) differ from the baseline category, and the contrasts between each level of political orientation and education on voter behavior.
